\documentclass{brandeis-dissertation3.12}

\usepackage[
backend=biber,
style=apa,
sorting=nyt
]{biblatex}

%\usepackage{amsmath}

\usepackage{graphicx}
\graphicspath{ {./mahmood_22_figures/}{./stone_2022_figs/}{./lin_2021_review_figs/} }
\usepackage{wrapfig}

\usepackage[hyphens,spaces,obeyspaces]{url}
\usepackage[colorlinks,allcolors=blue]{hyperref}

\addbibresource{mahmood_et_al.bib}
\addbibresource{stone_2022.bib}
\addbibresource{lin_2021_review.bib}
\addbibresource{intro_discussion_refs.bib}

\title{On Dynamic Neural Responses and Neural Network Interactions during Taste Processing}
\author{Abuzar Mahmood}
\graduationmonth{February}
\graduationyear{2022}
\program{Interdepartmental Program in Neuroscience}
\advisor{Donald B. Katz}
\signoff{Eric Chaslow}{Dean}
\committee{Stephen Van Hooser, Biology\\
Paul Miller, Biology\\
Ian Davison, Biology, Boston University}

\begin{document}

\maketitlepage
\makeapproval
\makecopyright

\begin{dissertation-acknowledgements}
I am indebted to many people in varying capacities for their guidance and support in helping me reach this point in my scientific career.

First, I would like to thank my advisor, Donald Brian Katz, for not only being very generous with his scientific acumen and wisdom, but also for actively making sure I enjoyed my PhD. Like many of those who have passed through your lab, Don, our interactions have left an (all but surely) indelible mark on me. I have deeply enjoyed not only our scientific, but also, philosophical and cultural discussions; and from those discussions, have not only learned valuable lessons about scientific concepts, but also about performing science. I am grateful to have been part of the vivacious and discerning environment you have fostered. 

I am also deeply grateful to my committee, Paul Miller, Steve Van Hooser, and Shantanu Jadhav. I am indebted to you for the generosity of your time, and the informality of your opinion. I have learned much not only from my committee meetings with you, but also from running into you up and down the hallways of Bassine and Volen. Learning from your diverse viewpoints and formidable combined neuroscience knowledge, has not only improved the direction of my work, but also has certainly made me a better scientist. I am truly thankful to have had you as mentors and colleagues.

I would like to thank our neighbors, the Jadhav, Van Hooser, and Miller labs for always being willing to lend a helping hand. In particular, I want to thank Ryan Young, and Ben Ballintyn, for always being willing to share their ideas and lend their expertise to guide me through my numerous technical problems.

It would be impossible to ignore my gratitude for my lab for being the rowdiest, kindest, most scientifically incisive bunch I have ever met. I have had the honor to be in the tutelage of many great mentors, who have indulged my scientific daydreams and guided my thoughts gently to sounder shores. I am particularly thankful to Dr. Jian-You Lin for being the General on the ground with clarity of vision and sound directions, Dr. Narendra Mukherjee for always challenging me to think through the core of my ideas and my analyses, Dr. Bradly T. Stone for holding my hand through many surgeries and being a voice of reason, Prof. Veronica Flores for exemplifying not only great science but also great lab member conduct, Daniel Svedberg for always being up for discussing ambitious and interesting ideas, Kathleen Maigler for being a model mentor from whose example I have learnt much and for being the glue that holds the lab together, and Max Bernstein for always being able to inject quirk into every task and discussion. I am also thankful to my mentees, Jessica Steindler, Hannah Germaine, and Thomas Murdy. It has been an honor working with you, and I hope I was able to teach you even a fraction of all I have learned from you. In particular, thank you Jess for literally carrying much of my project, and for being the ray of sunshine and fierce friend you are, mami.

I am grateful to the friends I have made in grad school and to my cohort for accompanying me on this journey. In particular, John, Rabia, Michael, Mohamed, Susannah, and Subhadra. I have been fortuitous to have had you to celebrate and commiserate, and have politically and morally charged discussions with you over these years. It is hard for me to decide whether I am a reflection of my circle, or vice-versa, but I have deeply cherished your unorthodox humor, and (as one may say politely) quirky outlook on life. I am thankful to Diane, Brandon, and Reign, for your warmth, thoughtfulness, and laughter; mostly Reign, but you already knew that. May we always witness you kick derriere, and may we always remain in your good graces (I shudder to think what would happen otherwise).

I am thankful to my undergraduate supervisor, Dr. Lakshmi Pulakat, who is single handedly responsible for planting the idea of pursuing a PhD in my head. The four years I spent in your lab were not only formative for me as a scientist, but also were deeply enjoyable and essential for molding my idea of what a great scientific environment should be like. I am thankful for your mentorship, your kindness, your support (in and out of science) over the years, and look forward to gleaning more wisdom from you in the future. I am also grateful to Madhavi Gavini for her friendship and for sharing her knowledge, experience, and perspective on multitudinous scientific and non-scientific conundrums over the years (from undergrad all through grad school).

I am deeply indebted to my family, without whom I would have never had the opportunity to apply to grad school. I have been blessed with five aunts and their equally caring husbands who have never let me feel far from home over the past decade, who have readily offered their time and effort, given sage advice whenever I needed it, and quite literally have fed me and clothed me. I am deeply indebted and immensely grateful for all you have done for me, which goes much beyond my graduate career. I want to thank my grandmother, my Nanoo, for her relentless support in my pursuit of this PhD, for her unwavering belief that I would succeed, and for the many, many, prayers she has made to my benefit. I am thankful to my mother, and my brothers, for always believing in me - I would not be here today if you hadn’t given me the option to choose my own path, and supported me as I walked on it.

Finally, I am thankful to my partner, Nishaat Mukadam, who has been a voice of reason through the ups and downs of my life through this PhD. I am grateful to have had your consultation and insight through the many challenges I have faced scientifically and non-scientifically, and I am lucky to have had both your iron-clad support and your infectious joy through these years. Thank you for your friendship, for teaching me so much over the years, and for never giving up on me. Merci pour tout.

\end{dissertation-acknowledgements}

\begin{dissertation-abstract}
The computational capacity of the brain (and artificial neural networks, which draw inspiration from brain function) is derived from the nonlinear transformations carried out in neural circuits. Such non-linearity is instantiated in the brain, not only by nonlinear transformation of neural “codes” across brain regions (which could be taken to correspond to layers in an artificial neural network), but also by similar dynamic (i.e. changing) transformations within single brain regions over time. Such temporal transformation of encoding in brain regions is allowed by recurrency of neural connections (circuits providing inputs to themselves) and generates dynamic activity patterns during neural processing for many tasks. Furthermore, such dynamics allow the same neurons to essentially behave in different ways over time, and hence perform different functions over time. Consequently, this means that neural circuits interact with each other in different ways over time. However, brain function is further influenced by (long-lasting) body states (e.g. sickness), context (e.g. new vs. familiar environments), and behavioral modes as attention vs. disengagement or awake vs. sleep. Therefore, “intrinsic” dynamics in the brain are occurring with a backdrop of changing external input, hence, to advance our understanding of neural processing we need to account for the intra- and inter-circuit interactions and dynamics that are embedded in a hierarchy of timescales. 

Although the dynamics of neural activity is a rapidly evolving topic of research, the contribution of this aspect of neural processing is still being understood. The work in this thesis investigates how neural processing and communication between neural populations change with brain/body states, and over shorter periods of time during stimulus-evoked responses in the context of taste processing. Starting with a general introduction into neural dynamics, and inter-region communication, I go on to discuss how the same neural population responds differently depending on body state (sickness). From there, I move on to discuss the dynamics of multi-region communication during taste processing, ending with an exploration of changing neural “encoding” over time, and discussing how any instance of neural responses can be considered as being generated from hierarchical processes at multiple timescales.
\end{dissertation-abstract}

\doublespacing

\tableofcontents

%\renewcommand{\listfigurename}{List of plots}
%\renewcommand{\listtablename}{Tables}
%\listoffigures
%\listoftables

\startbody

\include{chapter1.tex}
\include{chapter2.tex}
\include{chapter3.tex}
\include{chapter4.tex}
\include{chapter5.tex}

\end{document}